---
title: 'Exercises: Shiny Basics (htmlwidgets)'
output: pdf_document
---

```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- # ```{r, message=FALSE, warning=FALSE, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE) -->
<!-- # library(leaflet) -->
<!-- # ``` -->
<!-- #  -->
\begin{center}
\section{Exercises: Shiny and htmlwidgets}
\end{center}

# Overview of exercises

This collection of exercises is designed to introduce the basics of incorporating `htmlwidgets` into your shiny apps.


# Exercise 1: Shiny Africa Map

Previously you created a leaflet map that looks something like the image below, we're going to create a Shiny app with an embedded Leaflet map - which means you will need to load these libraries in RStudio.

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/leaflet-africa-data-plot.png")
grid.raster(img)
```

1.1 Create the shell of a shiny app from the template provided to you (a ui.R and server.R file)

1.2 Add the `leaflet` map generating code into the `server.R` file such that an output object will be created, that renders the map appropriately.

1.3 Display the map in the `ui.R` file using an appropriate function.

1.4 In the ui.R include the following controller:

- A slider that is allowed to move between 5 and 20, labelled "Standard Deviation"

1.5 Modify the `server.R` code such that changing the slider value changes the standard deviation of the `lat` and `lng` coordinates displayed, ensure that changing the slider affects the output map as you would expect.

1.6 Publish the shiny app to the web - ensure that it works!

# Exercise 2: Selected marker

For this exercise you will need to refer to [http://rstudio.github.io/leaflet/](http://rstudio.github.io/leaflet/)

2.1 In the `server.R` file create a new output UI element that will display the latitude and longitude coordinates of your selected points

2.2 Add appropriate code to the `ui.R` file such that this output object will be displayed.

2.3 Publish the shiny app to the web - ensure that it works!