---
title: "Exercises: Shiny Basics (Publishing)"
output: pdf_document
---
```{r, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- # ```{r, message=FALSE, warning=FALSE, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE) -->
<!-- # library(leaflet) -->
<!-- # ``` -->
<!-- #  -->
\begin{center}
\section{Exercises: Shiny Basics (Publishing)}
\end{center}

# Overview of Exercise

These exercises will take you through the steps necessary to connect your RStudio installation to your shinyapps.io account.

# Exercise 1

1.1 Sign up for an account at [www.shinyapps.io/admin/#/signup](http://www.shinyapps.io/admin/#/signup)

1.2 Choose an appropriate account name (this will be the "slug" for your shiny apps, i.e. http://fancyslug.shinyapps.io/myShinyApp)

1.3 Follow the steps on the website:

1.3.1 install.packages("rsconnect")

1.3.2 Copy your account info to the clipboard

1.3.3 Open RStudio's preferences, navigate to "Publishing" and click "connect"

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/connect_shinyapps-io.png")
 grid.raster(img)
```

1.3.4 Paste your account info into the dialog window.

1.4. If these steps have worked, you should be able to publish the shiny app you created in the previous exercise by clicking on the "Publish" button.

# Exercise 2

Login to your shinyapps.io and investigate how you might achieve the following:

- View overall active hours across all applications
- Modify the timeout period for your shiny app
- Archive your app
- Download a copy of the shiny app from the server




