---
title: "Exercises: Basic shinyApp and shinyapps.io"
output: pdf_document
---

```{r include=FALSE}
knitr::opts_chunk$set(fig.width=10, fig.height=6, fig.path='Figs/',
                      echo=FALSE, warning=FALSE, message=FALSE)
```

\begin{center}
\section{Exercises: Basic shinyApp and shinyapps.io}
\end{center}


# Overview of Exercises

These exercises take you through the steps to write shinyApps within your RMarkdown presentations.

# Exercise 1: Drawing Curves

1. Presentation setup:

1.1. Create a new project for these exercises - name it appropriately.

1.2. Create a new Slidy presentation file in your project

1.3. Change the first slide title to "Drawing Curves" and insert a new code chunk in which you'll write your shinyApp

1.4 Add `runtime: shiny` to the preamble of the presentation

2. Copy and paste the template for a Shiny app into your code chunk

```{r, echo=TRUE, eval=FALSE}
library(shiny)
shinyApp(
  ui = ,
  server = function(input, output){
    
  }
)
```

3. Add a fluidPage to the ui argument which contains the following arguments:

3.1 sliderInput with the following parameters:

- inputId = "exponent"
- label = "exponent"
- min = 1
- max = 5
- value 2

3.2 A wellPanel that contains text explaining what the slider is for

3.3 Knit your presentation together - you should see something like this

```{r, eval=F, echo=F}
library(shiny)
shinyApp(
  ui = fluidPage(
    wellPanel("This is a slider"),
    sliderInput("exponent", label = "exponent",
                min = 1,
                max = 5,
                value = 2)
  ),
  server = function(input, output){
    
  }
)
```

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/BasicShiny.png")
 grid.raster(img)
```

4. In the server function define a new variable for the output object (i.e. output$myThing) and assign it the following:

4.1 renderPlot

4.2 `curve(x^2, from = -5, to = 5)`

5. Add `plotOutput` to the ui argument and provide it the output object you defined above, knit together - you should have the following

```{r, eval=F, echo=F}
library(shiny)
shinyApp(
  ui = fluidPage(
    wellPanel("This is a slider"),
    sliderInput("exponent", label = "exponent",
                min = 1,
                max = 5,
                value = 2),
    plotOutput("myThing")
  ),
  server = function(input, output){
    output$myThing <- renderPlot({
      curve(x^2, from = -5, to = 5)
    })
  }
)
```

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/BasicShiny-2.png")
 grid.raster(img)
```

6. Modify the `curve` expression to be dependent on the input variable defined in `selectInput`

7. Knit together the presentation, the slider should update the plot now.

# Exercise 2: Registering a shinyapps.io account

Publishing the shiny enabled presentation requires a shinyapps.io account - this is easy to set up.

Knit the presentation file and select "publish" in the top-right of the window, you'll be presented with this dialog - select "Get started here"

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/Register-Shiny.png")
 grid.raster(img)
```

Register for an account - it's free. Choose an appropriate account name, this can be changed later but will require all your shiny apps to be re-deployed.

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/Shiny-account-name.png")
 grid.raster(img)
```

Navigate to Account -> Tokens and select "Show" at shinyapps.io to expose your secret token, copy and paste this into RStudio.

```{r fig.width=2, fig.height=10,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/Shiny-Token.png")
 grid.raster(img)
```

Provide the shiny app with a name (this will be the URL for the shiny app - it therefore cannot contain spaces or anything else that is *unsafe* in URLs) and press "Publish" - your app will then be deployed to the web and your browser opened when it's finished.


