---
title: "Exercises: Shiny Basics (Basic)"
output: pdf_document
---

```{r, message=FALSE, warning=FALSE, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

<!-- # ```{r, message=FALSE, warning=FALSE, include=FALSE} -->
<!-- # knitr::opts_chunk$set(echo = FALSE) -->
<!-- # library(leaflet) -->
<!-- # ``` -->
<!-- #  -->

\begin{center}
\section{Exercises: Shiny Basics}
\end{center}

# Overview of Exercises

This exercise is designed to consolidate the very fundamentals of a shiny app and the syntax required. They are minimal examples of what Shiny is capable of.

Use the same project for both exercises.

# Exercise 1

1.1 Create a project (refer back to the notes in the RMarkdown section of the course if necessary) containing both a ui.R and server.R file, you may find it useful to use the template files provided to you.

1.2 If you followed along with the instructors live-coding example, use the code from there. Otherwise you will need to add the following to your files:

ui:

```{r, eval=FALSE,echo=TRUE}
plotOutput("curvePlot")
```

server:

```{r, eval = FALSE, echo=TRUE}
output$curvePlot <- renderPlot({
    curve(x,
          from = -5,
          to = 5)
  })
```

1.3 Add a `sliderInput` to the ui that will allow users to change a value from 1 to 5, with a default value of 3. Refer to the documentation or http://shiny.rstudio.com/reference/shiny/latest/ as necessary.

1.4 Modify the server so that the curve(x^slider) is generated

1.5 Run the app and ensure it works.

# Exercise 2

2.1 Add a control to the Shiny app you built before that allows you to type your name

2.2 Use the "main" argument of `curve` to provide your name as the title of the plot

2.3 Use the paste function to combine the your name with the current value of the exponent, for instance: "Plot of x^4 (by Johanna Smith)"

```{r, eval=F,echo=TRUE}
paste("this", "that")
```

