---
title: "htmlwidgets"
output: 
  slidy_presentation
runtime: shiny
css: "www/teachingnotes.css"
---

# htmlwidgets: JavaScript

The interactive web is driven by JavaScript*, the majority of interactive elements that you use on websites are written in JavaScript - from interactive maps to auto-completing pop-up menus. 

```{r, echo=F}
library(shiny)
library(leaflet)
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
    sliderInput("latitutde", label = "Latitude", min = 5, max = 8, value = 6.9),
    sliderInput("longitude", label = "Longtidue", min = 15, max = 19, value = 17.7),
    sliderInput("sd", label = "Standard Deviation of Points", min = 5, max = 20, value = 10),
    selectInput("mapType", label = "Map Type:", choices = c("OpenStreetMap", "OpenStreetMap.BlackAndWhite","OpenTopoMap"))),
    mainPanel(
    leafletOutput("africa_map")
  ))),
  server = function(input, output) {
    
    output$africa_map <- renderLeaflet({
      
      africa_data_points = data.frame(
      lat = rnorm(100, mean = input$latitutde, sd = input$sd),
      lng = rnorm(100, mean = input$longitude, sd = input$sd),
      size = runif(100, 5, 10),
      popup = as.character(1:100)
    )
    
      m <- leaflet(africa_data_points) %>%
        addProviderTiles(input$mapType)
      m %>% addCircleMarkers(radius = ~size, popup = ~popup)
      
    })
  },options = list(height = "100%")
)

```

# htmlwidgets: The Framework

Like in R, there are hundreds of different JavaScript libraries dedicated to various visualisation tasks. 

There is a package called `htmlwidgets` that allows R developers to easily build bindings to JavaScript libraries, allowing incredibly rich interactive content for the web to be built just with the R language.

# htmlwidgets: The Packages

These bindings to JavaScript libraries are typically distributed as individual R packages; an individual R package for an individual JavaScript library. The <a href="http://htmlwidgets.org">htmlwidgets.org</a> website provides a showcase of some of the `htmlwidget` dependent bindings that are available through CRAN.

# leaflet maps

Let's make an interactive map together using the `leaflet` library:

```{r}
library(leaflet)
leaflet() %>%
  addTiles()
```


# What is %>%?

Every single* htmlwidget library uses `%>%` [the pipe operator] as it makes it very easy to build up expressions/operations/visualisations.

For an example of why we want to use pipes, I'll borrow the [following example](https://twitter.com/AmeliaMN/status/748193609401327616) from Hadley Wickham:

<div class="incremental">  
> Little bunny Foo Foo </br>
> Went hopping through the forest </br>
> Scooping up the field mice </br>
> And bopping them on the head </br>
</div>

<div class="incremental">  
```{r, eval=FALSE}
foo_foo <- little_bunny()
```
</div>

<div class="incremental">  
```{r, eval=FALSE}
bop_on(
  scoop_up(
    hop_through(foo_foo, forest),
    field_mouse
  ),
  head
)
```
</div>

<div class="incremental">  
```{r, eval=FALSE}
foo_foo %>%
  hop_through(forest) %>%
  scoop_up(field_mouse) %>%
  bop_on(head)
```
</div>

# Simpler %>% example

The pipe operator takes the left-hand side of your expression and inserts it into the first argument of the right-hand side of the expression:

<div class="incremental">  
```{r, eval = FALSE}
"cats" %>% rep(4)
```
</div>

<div class="incremental">  
Almost completely beyond the scope of this course, the left-hand side can be inserted elsewhere in the right-hand side of the expression using `.`
</div>

<div class="incremental"> 
```{r}
"cats" %>% paste("too","many", .)
```
</div>
<p>
<div class="incremental"> 
The `%>%` operator comes from the `magrittr` library and can be read about more [here](https://github.com/smbache/magrittr). While it is often incredibly useful, it is not a hammer to be used without exception - some code is both harder to write and code with pipes.
</div>

<div class="incremental"> 
Also - you're allowed to hate pipes. That's fine.
</div>

# Back to leaflet

`leaflet()` instantiates a map, within which we can put data. `addTitles` loads the default "tiles" for the map

<div class="incremental"> 
```{r}
africa_data_points = data.frame(
  lat = rnorm(26, mean = 6.9, sd = 10),
  lng = rnorm(26, mean = 17.7, sd = 10),
  size = runif(26, 5, 10),
  label = letters
  )
leaflet(data = africa_data_points) %>%
  addTiles()
```

`addCircleMarkers` intelligently looks inside of the `data` for columns that might be longitude and latitude coordinates and creates circles on top of these points.

```{r}
leaflet(data = africa_data_points) %>%
  addTiles() %>%
  addCircleMarkers()
```
</div>

# Plotly and formulae

[plot.ly](http://plot.ly) is both a JavaScript library capable of generating interactive charts & maps and a "point-and-click" tool for building interactive visualisations in the web browser. 

It's highly recommended you check it out, as most of the functionality of the htmlwidget library can be figured out from building charts interactively in the browser.

It's also a good way to introduce the `~column` syntax.

# Plotly and formulae

Let's use sample data from the `WDI` library that allows us to access the World Data Bank:

```{r, message=FALSE, warning=FALSE}
library(WDI)
world_data <-
  WDI(
  country = c("GB", "EG", "SA", "EE", "CA"),
  indicator = c("EG.GDP.PUSE.KO.PP","EN.ATM.CO2E.PP.GD.KD","NY.GDP.MKTP.KD"),
  start = 1990,
  end = 2000
  )
str(world_data)
```

<div class="incremental"> 

The basic `plot_ly` chart can be constructed as follows, `~` allows us to access the "environment" of the `data` argument. 

The `~` is more usually used in formulae, for instance `pop ~ birth_rate`.

```{r, message=FALSE, warning=FALSE}
library(plotly)
plot_ly(data = world_data,
        x = ~year,
        y = ~EG.GDP.PUSE.KO.PP,
        color = ~country)
```

Full documentation of the API is available here: [plot.ly/r/reference/](https://plot.ly/r/reference/)

```{r, message=FALSE, warning=FALSE}
plot_ly(data = world_data,
        x = ~year,
        y = ~EG.GDP.PUSE.KO.PP,
        color = ~country,
        type = "scatter",
        mode = "lines") %>%
  layout(xaxis = list(title = "Year"))
```
</div>

# Using htmlwidgets

Pipes can become confusing and are not always useful, particularly if one of the steps in your chain relies on a conditional. 

It's often a good idea to assign intermediate stages in your visualisation:

<div class="incremental"> 
```{r}
basic_plot <- plot_ly(
  data = world_data,
  x = ~ year,
  y = ~ EG.GDP.PUSE.KO.PP,
  color = ~ country,
  type = "scatter",
  mode = "lines"
) %>%
  layout(xaxis = list(title = "Year"))
```

```{r}
label_length <- "long"
```

```{r}
if (label_length == "short") {
  basic_plot %>%
    layout(xaxis = list(title = "EG.GDP.PUSE.KO.PP"))
} else {
  basic_plot %>%
    layout(xaxis = list(title = "GDP per unit of energy use (PPP $ per kg of oil equivalent)"))
}
```
</div>

# Exercises (30mins)

That's a very, very quick and basic introduction to two `htmlwidget` libraries it's very important to understand they're simulataneously very different and quite similar to one another.

In completing these exercises please do refer to the documentation for the appropriate library at [htmlwidgets.org](http://www.htmlwidgets.org)

# Choosing htmlwidgets and tutorials

There are more than 15 htmlwidget libraries available, many have overlapping features but others are very unique (see https://github.com/jbkunst/rchess).

How do you decide to which library to choose, an what visualisation is most appropriate for your data?

<div class="incremental"> 
The following guide is currently under development [ox-it.github.io/OxfordIDN_htmlwidgets](http://ox-it.github.io/OxfordIDN_htmlwidgets) and aims to answer the following questions:

- Which library is capable of making chart X?
- Which charts can be made with library X?
- What type of data can be displayed with each chart/library?
- How does the process for creating chart X compare across the available libraries?
</div>

<div class="incremental">
There is also a very alpha library called `oidnChaRts` that aims to make it simple to switch between different htmlwidget libraries.
</div>



