---
title: "Visualising-Geographic-Data"
output: 
  slidy_presentation
runtime: shiny
css: "www/teachingnotes.css"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# Scattergeo plots

The simplest and most versatile demo of visualising geographic data is a scattergeo plot using leaflet.

We're follow these steps:

- Import data
- Filter data
- Visualise data

# Importing Data with readr

The `readr` library is your recommended library for importing rectangular datasets, for instance a set of longitude-latitude pairs.

From the downloaded materials, find the following file:

training_geo-intro-to-R/Handouts/capitals_with_locations.csv

Create a new folder in your project called "data"

Copy the file there

# read_csv and data.frames

The data can now be imported with the following command:

```{r}
capitals_locations <- read_csv("data/capitals_with_locations.csv")
```

# Filtering data with dplyr

The `dplyr` library provides an excellent framework for filtering, sorting, augmenting and simplfying datasets. Remember there is an excellent cheatsheet on dplyr from Help -> Cheatsheets.

Assume we're interested in removing portions of the data:

<div class="incremental"> 
```{r}
library(dplyr)
capitals_locations %>%
  filter(!Continent == "EU")
```
</div>


# What is %>%?

Almost all libraries from the tidyverse use `%>%` [the pipe operator] as it makes it very easy to build up expressions or operations. It is also incredibly popular in the htmlwidget libraries we'll soon cover.

For an example of why we want to use pipes, I'll borrow the [following example](https://twitter.com/AmeliaMN/status/748193609401327616) from Hadley Wickham:

<div class="incremental">  
> Little bunny Foo Foo </br>
> Went hopping through the forest </br>
> Scooping up the field mice </br>
> And bopping them on the head </br>
</div>

<div class="incremental">  
```{r, eval=FALSE}
foo_foo <- little_bunny()
```
</div>

<div class="incremental">  
```{r, eval=FALSE}
bop_on(
  scoop_up(
    hop_through(foo_foo, forest),
    field_mouse
  ),
  head
)
```
</div>

<div class="incremental">  
```{r, eval=FALSE}
foo_foo %>%
  hop_through(forest) %>%
  scoop_up(field_mouse) %>%
  bop_on(head)
```
</div>

# Simpler %>% example

The pipe operator takes the left-hand side of your expression and inserts it into the first argument of the right-hand side of the expression:

<div class="incremental">  
```{r, eval = FALSE}
"cats" %>% rep(4)
```
</div>

<div class="incremental">  
Almost completely beyond the scope of this course, the left-hand side can be inserted elsewhere in the right-hand side of the expression using `.`
</div>

<div class="incremental"> 
```{r}
"cats" %>% paste("too","many", .)
```
</div>
<p>
<div class="incremental"> 
The `%>%` operator comes from the `magrittr` library and can be read about more [here](https://github.com/smbache/magrittr). While it is often incredibly useful, it is not a hammer to be used without exception - some code is both harder to write and code with pipes.
</div>

<div class="incremental"> 
Also - you're allowed to hate pipes. That's fine.
</div>

# Back to dplyr

We could combine multiple operations from the `dplyr` library together:

```{r}
capitals_locations %>%
  filter(!Continent == "EU") %>%
  filter(country.population < 10000) %>%
  select(Country, Capital, country.population, longitude, latitude)
```

# Visualising this data in R

The `ggplot` is an extremely popular and powerful tool for creating static visualisations in R.

```{r}
library(rworldmap)
library(ggplot2)
map.world <- map_data(map="world")

ggplot() + 
  geom_map(data=map.world, map=map.world, aes(map_id=region, x=long, y=lat)) + 
  geom_point(data = capitals_locations, aes(x = longitude, y = latitude), color = "red")
```


# htmlwidgets: JavaScript

The interactive web is driven by JavaScript*, the majority of interactive elements that you use on websites are written in JavaScript - from interactive maps to auto-completing pop-up menus. 

```{r, echo=F}
library(shiny)
library(leaflet)
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
    sliderInput("latitutde", label = "Latitude", min = 5, max = 8, value = 6.9),
    sliderInput("longitude", label = "Longtidue", min = 15, max = 19, value = 17.7),
    sliderInput("sd", label = "Standard Deviation of Points", min = 5, max = 20, value = 10),
    selectInput("mapType", label = "Map Type:", choices = c("OpenStreetMap", "OpenStreetMap.BlackAndWhite","OpenTopoMap"))),
    mainPanel(
    leafletOutput("africa_map")
  ))),
  server = function(input, output) {
    
    output$africa_map <- renderLeaflet({
      
      africa_data_points = data.frame(
      lat = rnorm(100, mean = input$latitutde, sd = input$sd),
      lng = rnorm(100, mean = input$longitude, sd = input$sd),
      size = runif(100, 5, 10),
      popup = as.character(1:100)
    )
    
      m <- leaflet(africa_data_points) %>%
        addProviderTiles(input$mapType)
      m %>% addCircleMarkers(radius = ~size, popup = ~popup)
      
    })
  },options = list(height = "100%")
)

```

# htmlwidgets: The Framework

Like in R, there are hundreds of different JavaScript libraries dedicated to various visualisation tasks. 

There is a package called `htmlwidgets` that allows R developers to easily build bindings to JavaScript libraries, allowing incredibly rich interactive content for the web to be built just with the R language.

# htmlwidgets: The Packages

These bindings to JavaScript libraries are typically distributed as individual R packages; an individual R package for an individual JavaScript library. The <a href="http://htmlwidgets.org">htmlwidgets.org</a> website provides a showcase of some of the `htmlwidget` dependent bindings that are available through CRAN.

# Introduction to leaflet

`leaflet()` instantiates a map, within which we can put data. `addTitles` loads the default "tiles" for the map

<div class="incremental"> 
```{r}
library(leaflet)
capitals_locations %>%
leaflet() %>%
  addTiles()
```

`addCircleMarkers` intelligently looks inside of the `data` for columns that might be longitude and latitude coordinates and creates circles on top of these points.

```{r}
capitals_locations %>%
leaflet() %>%
  addTiles() %>%
  addCircleMarkers()
```
</div>

# Plotly

[plot.ly](http://plot.ly) is both a JavaScript library capable of generating interactive charts & maps and a "point-and-click" tool for building interactive visualisations in the web browser. 

It's highly recommended you check it out, as most of the functionality of the htmlwidget library can be figured out from building charts interactively in the browser.

It's also a good way to introduce the `~column` syntax.

# Plotly and formulae

Let's use sample data from the `WDI` library that allows us to access the World Data Bank:

```{r, message=FALSE, warning=FALSE}
library(WDI)
world_data <-
  WDI(
  country = c("GB", "EG", "SA", "EE", "CA"),
  indicator = c("EG.GDP.PUSE.KO.PP","EN.ATM.CO2E.PP.GD.KD","NY.GDP.MKTP.KD"),
  start = 1990,
  end = 2000
  )
str(world_data)
```

<div class="incremental"> 

The basic `plot_ly` chart can be constructed as follows, `~` allows us to access the "environment" of the `data` argument. 

The `~` is more usually used in formulae, for instance `pop ~ birth_rate`.

```{r, message=FALSE, warning=FALSE}
library(plotly)
plot_ly(data = world_data,
        x = ~year,
        y = ~EG.GDP.PUSE.KO.PP,
        color = ~country)
```

Full documentation of the API is available here: [plot.ly/r/reference/](https://plot.ly/r/reference/)

```{r, message=FALSE, warning=FALSE}
plot_ly(data = world_data,
        x = ~year,
        y = ~EG.GDP.PUSE.KO.PP,
        color = ~country,
        type = "scatter",
        mode = "lines") %>%
  layout(xaxis = list(title = "Year"))
```
</div>


# Exercises (30mins)

That's a very, very quick and basic introduction to two `htmlwidget` libraries it's very important to understand they're simulataneously very different and quite similar to one another.

In completing these exercises please do refer to the documentation for the appropriate library at [htmlwidgets.org](http://www.htmlwidgets.org)












