---
title: "Adding More Interactivity to Shiny"
output: pdf_document
css: "www/printingnotes.css"
---

```{r, echo=F, message=FALSE, warning=FALSE}
suppressWarnings(require(leaflet))
```


## Warnings about Shiny

- Controller variables can only be used once in a shiny app - `sliderInput("thisVar", ...)` only one slider can be created in the app that uses `input$thisVar`

- Duplicate output cannot be used in shinyApps, it will create errors. So you cannot have multiple instances of `plotOutput("myFancyPlot")` in your app

## Interactive Data Visualisations

So far we've built a Shiny app with interactive elements, but not interactive charts. But *all* htmlwidget-dependent interactive elements can be inserted into Shiny apps.

Many of the htmlwidget-dependent libraries have built in Shiny integration - allowing actions (like clicking) inside of a chart to be communicated to the Shiny server code.

## Leaflet in Shiny

The following Shiny app uses Leaflet:

```{r, echo = FALSE, eval=FALSE}
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
    sliderInput("latitutde", label = "Latitude", min = 5, max = 8, value = 6.9),
    sliderInput("longitude", label = "Longtidue", min = 15, max = 19, value = 17.7),
    sliderInput("sd", label = "Standard Deviation of Points", min = 5, max = 20, value = 10),
    selectInput("mapType", label = "Map Type:", choices = c("OpenStreetMap", "OpenStreetMap.BlackAndWhite","OpenTopoMap"))
    ),
    mainPanel(
    leafletOutput("africa_map")
  ))),
  server = function(input, output) {
    
    africa_data_points <- reactive({data.frame(
      lat = rnorm(100, mean = input$latitutde, sd = input$sd),
      lng = rnorm(100, mean = input$longitude, sd = input$sd),
      size = runif(100, 5, 10),
      popup = as.character(1:100)
    )})
    
    output$africa_map <- renderLeaflet({
      
      africa_data_points <- africa_data_points()
    
      m <- leaflet(africa_data_points) %>%
        addProviderTiles(input$mapType)
      m %>% addCircleMarkers(radius = ~size, popup = ~popup)
      
    })
  },options = list(height = "700px")
)
```


```{r, echo=F}
library(png)
library(grid)
img <- readPNG("images/htmlwidgets_africa_map.png")
 grid.raster(img)
```


## Reactivity

Building interactive elements with Shiny is incredibly simple, the difficulty arises in controlling that interactivity - when and how things update.

In the Shiny framework interactivity is achieved through `reactive` expressions which essentially invalidate one another, by moving a slider you are updating a variable which invalidates the dependent output resulting in re-evaluation.

## Reactivity

The simplest demonstration of reactive expressions is provided by this example:

```{r, eval=F, echo=F}
shinyApp(
  ui = fluidPage(
    sliderInput("a", label = "a", min = 1, max = 5, value = 2),
    textInput("b", label = "b", value = "x axis"),
    plotOutput("hist")
  ),
  
  server = function(input, output){
    
    output$hist <- renderPlot(
      hist(rnorm(1000, mean = input$a), xlab = input$b)
    )
    
  }
)
```

```{r, echo=F}
library(png)
library(grid)
img <- readPNG("images/htmlwidgets_africa_map.png")
 grid.raster(img)
```
