---
title: "Adding Shiny to Presentations"
output: pdf_document
css: "www/printingnotes.css"
---

## Shiny and shinyapps.io

Shiny is an R package that allows interactive HTML elements to be generated (and *rendered*) from interpreted R code.

To display Shiny content in a webpage the R code must be interpreted by a Shiny-enabled server - [shinyapps.io](http://www.shinyapps.io) provides this as a service.

However it is very simply to interact with Shiny apps on your local machine without relying on such services. In the exercises you'll create your own shinyapps.io account and upload a Shiny app to the web, but for the next few slides we'll consider only what you need to do to build an interactive Shiny app on your local machine.

## First Shiny App

To use shiny it's necessary to install the shiny library as follows, and make sure that it is "loaded" using the `library` function

```{r, show_installing_shiny, echo=TRUE, eval=FALSE}
install.packages("shiny")
library(shiny)
```

The following interactive app will now be built by your lecturer and the mechanics of it explained.

```{r, exponent_curve_example, echo=FALSE, eval=FALSE}
shinyApp(
  ui = fluidPage(
    sliderInput("exponent", min = 1, max = 5, step = 1, value = 1, label = "Exponent"),
    plotOutput("curvePlot")
  ),
  server = function(input, output){
    output$curvePlot <- renderPlot({
      curve(x^input$exponent, from = -5, to = 5)
    })
  }
)
```


```{r fig.width=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/curve-shiny-app.png")
 grid.raster(img)
```


## Why UI and Server?

Shiny apps are split into two sections - the UI and server side code.

It's important to understand what Shiny does - it provides a way to write a HTML GUI (i.e. HTML and JavaScript) in R code and provides a framework for a browser displaying a Shiny app to exchange information with a Shiny server.

The app is therefore split into client-side and server-side code - or shinyUI and shinyServer. Fundamentally, the UI is only aware of "values" (or data) from the server assigned to the output object.

## Exercises (15mins)

## Bootstrap and Shiny apps

[Bootstrap](http://getbootstrap.com/) is an incredibly powerful and flexible framework for making "responsive web elements" - content that resizes (or transforms) dependent on the window size of your browser.

The following interface is built entirely within Shiny using the following layout tools:

- navbarPage

- navbarMenu

- tabsetPanel

```{r, bootstrap_shiny, echo=FALSE, eval=FALSE}
library(shiny)
shinyApp(
  ui = navbarPage(
    "navbarMenu Example",
    tabPanel("fluidPage",
             fluidPage(
               sliderInput("number.of.data", label = "Number of Data", min = 200, max = 1000, value = 500),
               plotOutput("hist.1")
             ))
    ,
    navbarMenu("navbarMenu",
                tabPanel("tabsetPanel", 
                         fluidPage(
                           sidebarLayout(
                             sidebarPanel(sliderInput("number.of.data", label = "Number of Data", min = 200, max = 1000, value = 500)),
                             mainPanel(tabsetPanel(
                               tabPanel("Histogram",
                                        plotOutput("hist.2")),
                               tabPanel("Other",
                                        "Other")
                             ))
                           )
                         )
                         ),
                tabPanel("Child 2", "foobar")),
    collapsible = TRUE
  ),
  server = function(input, output){
    
    data.for.histogram <- reactive({
      rnorm(n = input$number.of.data, mean = 0, sd = 2)
    })
    
    output$hist.1 <- renderPlot({
      
      data <- data.for.histogram()
      
      hist(data)
    })
    
     output$hist.2 <- renderPlot({
      
      data <- data.for.histogram()
      
      hist(data)
    })
  }, options = list(width = "100%", height = "800px")
)
```

```{r fig.width=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/navbarPage-Shiny.png")
 grid.raster(img)
```


## Exercises (10mins)




