---
title: "htmlwidgets"
output: pdf_document
css: "www/printingnotes.css"
---

## htmlwidgets: JavaScript

The interactive web is driven by JavaScript*, the majority of interactive elements that you use on websites are written in JavaScript - from interactive maps to auto-completing pop-up menus. 

```{r, fig.height=10, echo=F, eval=F}
library(shiny)
library(leaflet)
shinyApp(
  ui = fluidPage(
    sidebarLayout(
      sidebarPanel(
    sliderInput("latitutde", label = "Latitude", min = 5, max = 8, value = 6.9),
    sliderInput("longitude", label = "Longtidue", min = 15, max = 19, value = 17.7),
    sliderInput("sd", label = "Standard Deviation of Points", min = 5, max = 20, value = 10),
    selectInput("mapType", label = "Map Type:", choices = c("OpenStreetMap", "OpenStreetMap.BlackAndWhite","OpenTopoMap"))),
    mainPanel(
    leafletOutput("africa_map")
  ))),
  server = function(input, output) {
    
    output$africa_map <- renderLeaflet({
      
      africa_data_points = data.frame(
      lat = rnorm(100, mean = input$latitutde, sd = input$sd),
      lng = rnorm(100, mean = input$longitude, sd = input$sd),
      size = runif(100, 5, 10),
      popup = as.character(1:100)
    )
    
      m <- leaflet(africa_data_points) %>%
        addProviderTiles(input$mapType)
      m %>% addCircleMarkers(radius = ~size, popup = ~popup)
      
    })
  },options = list(height = "100%")
)

```


```{r fig.width=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/htmlwidgets_africa_map.png")
 grid.raster(img)
```

## htmlwidgets: The Framework

Like in R, there are hundreds of different JavaScript libraries dedicated to various visualisation tasks. 

<p>

There is a package called `htmlwidgets` that allows R developers to easily build bindings to JavaScript libraries, allowing incredibly rich interactive content for the web to be built just with the R language.

## htmlwidgets: The Packages

These bindings to JavaScript libraries are typically distributed as individual R packages; an individual R package for an individual JavaScript library. The <a href="http://htmlwidgets.org">htmlwidgets.org</a> website provides a showcase of some of the `htmlwidget` dependent bindings that are available through CRAN.

```{r, echo=T, eval = F}
m <- leaflet() %>%
  addTiles() %>%  # Add default OpenStreetMap map tiles
  addMarkers(lng=174.768, lat=-36.852, popup="The birthplace of R")
m
```

```{r fig.width=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/htmlwidgets_africa_map.png")
 grid.raster(img)
```

```{r, message=FALSE, warning=FALSE, eval=F}
library(visNetwork)
suppressWarnings(library(igraph))
igraph_network <- graph.famous("Walther")
data <- toVisNetworkData(igraph_network)
visNetwork(nodes = data$nodes, edges = data$edges) %>%
  visNodes(color = list(background = "red", 
                        border = "darkblue",
                        highlight = "yellow")) %>%
  visEdges(color = list(color = "red"))
```


```{r fig.width=4,echo=FALSE}
library(png)
library(grid)
img <- readPNG("images/htmlwidgets_visNetwork.png")
 grid.raster(img)
```


## Exercises (20mins)