---
title: "Stand Alone Shiny Apps"
output: pdf_document
css: "www/printingnotes.css"
---

## Split File Shiny Apps

In this course we have only considered how to build Shiny apps directly within an RMarkdown chunk, however this is only really applicable for small-scale Shiny apps.

These apps we've been building could be considered "self-contained" as they are written in a single call to the `shinyApp` function, large scale apps are developed across at least two files:

- ui.R

- server.R

This "split-file" Shiny app setup allows for much greater "separation of concerns" and make development (and reuse) of an application easier. 

## Embedding "Split File" Shiny Apps into Presentations

There are two methods of emebedding split file apps into RMarkdown presentations:

>- `shinyAppDir`: requires the entire RMarkdown document to be in the shiny runtime, i.e. shinyapps.io active hours are consumed while anywhere in the presentation

>- iframes: allows a shinyapps.io hosted Shiny app to be embedded into a single (or many) slides without the rest of the app requiring Shiny active hours to display

## Layout of Split Files

The ui.R and server.R files for a split file Shiny App must be contained within the same directory, and have the following structure

```{r, echo=TRUE, eval=FALSE}
## ui.R
library(plotly) # You must load libraries into the ui.R file when they are used in client-side interactions; this includes all libraries for interactive charts
shinyUI(
  fluidPage(
    ...,
  )
)
```

```{r, echo=TRUE, eval=FALSE}
## server.R
library(plotly) # You must load libraries into the server.R file when they are used in server-side interactions; in general every package your app is dependent on must be loaded here
shinyServer(function(input, output){
  ...,
  })
```

## Loading local files into split file apps

This is largely beyond the scope of this course, however it is useful to have some advisory rules:

- CSS and JavaScript should be kept in the `www` sub-directory of the app
- Images should be kept in the `images` sub-directory of the app
- All paths must be given relative to the ui.R and server.R folder (the root directory of the app)

Note that ALL files in the Shiny app directory will be uploaded to shinyapps.io if you choose to deploy your app there.